Geometric models reveal behavioral and neural signatures of how naturalistic experiences are transformed into episodic memories

### 2020-08-19 paper/main.tex
JRM inputted handwritten edits

\begin{abstract}
The ways in which our experiences unfold over time are, by nature, incalculably complex.  One effective approach to making sense of this complexity is to characterize experiencess as unique \textit{trajectories} through a certain geometric representational space.  In word embedding spaces, for example, each dimension reflects a concept, such that a given coordinate reflects a weighted mixture of those concepts.  We propose a framework for projecting naturalistic experiences into word embedding spaces, such that the conceptual content of each moment of an experience, and how different moments of the experience relate, are reflected by the \textit{shape} of the experience's trajectory.  By projecting memories of those experiences into the same spaces, one may then geometrically compare the shape of the original experience's trajectory to the shape of how it is remembered later.  According to this view, encoding an experience into memory entails geometrically distorting or transforming the original experience's trajectory.  This translates qualitative questions about how we remember naturalistic experiences into quantitative geometric comparisons.  We applied our framework to data collected as participants watched and verbally recounted a television episode while undergoing functional neuroimaging.  We found that the trajectory of each participant's recall reflected the high-level \textit{essence} (i.e., large-scale narrative summaries) of the episode's trajectory, but participants differed markedly in their memories for low-level features (i.e, small-scale details).  We also identified a network of brain structures that were sensitive to the shape of the episode's trajectory through word embedding space, and an overlapping network that predicted, at the time of encoding, how people would distort (transform) the episode's trajectory when they recounted the episode later.  Our work provides insights into how our brains transform ongoing experiences when we encode them into episodic memories, and provides a formal geometric framework for characterizing the complex dynamic content of naturalistic experiences.
\end{abstract}

@@

Statistical models of memory studies often treat recall success as binary (in other words, an item either was or was not recalled), or occasionally categorical \citep[e.g., to distinguish familiarity from recollection;][]{YoneEtal02}.  Such approaches are tenable in classical list-learning or recognition memory paradigms, as the presented stimuli tend to be very simple (e.g., a sequence of individual words or items).  However, memory for naturalistic experiences is much more nuanced.  For example, certain aspects of an experience might be correctly remembered at varying levels of detail, or distorted, or forgotten entirely.  Further, each remembering is itself a richly structured phenomenon.  Our framework produces a content-based model of individual video and recall events by projecting the dynamic content of the video and participants' recalls into a shared topic space.  This allows for direct, quantitative comparisons between all stimulus and recall events, as well as between the recall events themselves.  Leveraging these content-based models of the stimulus/recall events, we developed two novel, \textit{continuous} metrics for analyzing naturalistic memory:~\textit{precision} and \textit{distinctiveness}.  Precision is intended to capture the ``completeness'' of recall, or how fully the presented content was recapitulated in memory.  We define a recall event's precision as the maximum correlation between the topic proportions of that recall event and any video event (Fig.~\ref{fig:precision-distinctiveness}).  A second novel metric we introduce here is \textit{distinctiveness}, which is intended to capture the ``specificity'' of recall.  In other words, distinctiveness quantifies the extent to which a given recalled event reflects the most similar presented event moreso than it does other presented events.  To compute a recall event's distinctiveness, we first identify the video event to which its topic vector is most strongly correlated.  We then define distinctiveness as one minus the average correlation between the given recall event and all \textit{other} video events.  In addition to individual events, one may also use these metrics to describe each participant's overall performance by averaging across a participant's event-wise precision or distinctiveness scores.  Participants whose recall events are more veridical descriptions of what happened in the video event will presumably have higher precision scores. We find that, across participants, higher precision scores are positively correlated with both hand-annotated memory performance (as collected by Chen et al., 2017; Pearson's $r(15) = 0.60, p = 0.010$) and the number of video events successfully remembered, as determined by our model (Pearson's $r(15) = 0.90, p < 0.001$).  We also hypothesized that participants who recounted events in a more distinctive way would display better overall memory.  We find that participants' distinctiveness scores were
correlated with our model's estimated number of recall events (Pearson's $r(15) = 0.55, p = 0.021$).  However, we found no evidence that distinctiveness scores were correlated with hand-annotated memory performance (Pearson's $r(15) = 0.28, p = 0.275$).  We elaborate on this potential discrepency in the \textit{Discussion} section.

%% JRM STOPPED HERE

@@

Further intuition for the behaviors captured by these two metrics may be gained by directly examining the content of the video and recalls our framework models.  In Figure \ref{fig:precision-detail}, we contrast recalls for the same video event (event 22) from two participants: one with a high precision score (P17), the other with a low precision score (P6).  From the HMM-identified event boundaries, we recovered the set of annotations describing the content of an example video event (Fig.~\ref{fig:precision-detail}B), and divided them into different color-coded sections for each action or feature described.  We then similarly recovered the set of sentences comprising the corresponding recall event for each of the two example participants.  Because the recall sliding windows overlap heavily, and each recall event spans multiple recall timepoints (i.e., windows), we have stripped any sentences from the beginning and end that describe earlier or later video events for the sake of readability.  In other words, Fig.~\ref{fig:precision-detail}C shows a subset of the full recall event text, comprising sentences between the first and last descriptions of content from the example video event.  We then colored all words describing actions and features coded in panel B by their corresponding color.  Visual comparison of the transcripts reveals that the most precise participant's recall both captures more of the video event's content, and does so with far more detail.

Figure \ref{fig:distinctiveness-detail} similarly contrasts two example participants' recalls for a common video event (event 19) to illustrate the tangible differences between high and low distinctiveness scores.  Here, we have extracted the full set of sentences comprising the most distinctive recall event (P13) and least distinctive recall event (P11) matched to the example video event (Fig.~\ref{fig:distinctiveness-detail}C).  We also extracted the annotations for the example video event, as well as those from each other video event whose content the example participants' single recall events described (Fig.~\ref{fig:distinctiveness-detail}B).  We then shaded the annotation text for each video event with a different color, and shaded each word of the example participants' recall text by the color of the video event it describes.  The majority of the most distinctive recall event text describes video event 19's content, with the first five and last one sentence describing the video events immediately preceding and succeeding the current one, respectively.  In contrast, the least precise participant's recall for video event 19 blends the content from five separate video events, does not transition between them in order, and often combines descriptions of two video events' content in the same sentence.

@@

In addition to the more ``holistic'' measure of memory described in the previous section, our framework also affords the ability to drill down to individual words and quantify how each word relates to the  memorability of each event. The results displayed in Figures \ref{fig:list-learning}C and \ref{fig:precision-detail}A suggest that certain events were remembered better than others.  Given this, we next asked asked whether the events were generally remembered well or poorly tended to reflect particular content.  Because our analysis framework projects the dynamic video content and participants' recalls into a shared space, and because the dimensions of that space represent topics (which are, in turn, sets of weights over words in the vocabulary), we are able to recover the weighted combination of words that make up any point (i.e., topic vector) in this space.  We first computed the average precision with which participants recalled each of the 30 video events (Fig.~\ref{fig:topics}A; note that this result is analogous to a serial position curve created from our continuous recall quality metric).  We then computed a weighted average of the topic vectors for each video event, where the weights reflected how reliably each event was recalled.  To visualize the result, we created a ``wordle'' image~\citep{MuelEtal18} where words weighted more heavily by better-remembered topics appear in a larger font (Fig.~\ref{fig:topics}B, green box).  Across the full video, content that reflected topics necessary to convey the central focus of the video (e.g., the names of the two main characters, ``Sherlock" and ``John", and the address of a major recurring location, ``221B Baker Street") were best remembered.  An analogous analysis revealed which themes were poorly remembered.  Here in computing the weighted average over events' topic vectors, we weighted each event in \textit{inverse} proportion to how well it was remembered (Fig.~\ref{fig:topics}B, red box).  The least well-remembered video content reflected information not necessary to later convey a general summary of the video, such as the proper names of relatively minor characters (e.g., ``Mike," ``Molly," and ``Lestrade") and locations (e.g., ``St. Bartholomew's Hospital").
%, as well as the brief, animated clip participants viewed at the beginning of each of the two scan session (involving ``singing" ``cartoon" characters).

@@

A similar result emerged from assessing the topic vectors for individual video and recall events (Fig.~\ref{fig:topics}C).  Here, for each of the three best- and worst-remembered video events, we have constructed two wordles: one from the original video event's topic vector (left) and a second from the average recall topic vector for that event (right).  The three best-remembered events (circled in green) correspond to scenes important to the central plot-line: a mysterious figure spying on John in a phone booth; John meeting Sherlock at Baker St. to discuss the murders; and Sherlock laying a trap to catch the killer.  Meanwhile, the three worst-remembered events (circled in red) reflect scenes that are non-essential to summarizing the narrative's structure: the video of singing cartoon characters participants viewed prior to the main episode; John asking Molly about Sherlock's habit of over-analyzing people; and Sherlock noticing evidence of Anderson's and Donovan's affair.

@@

While a large number of language models exist \citep[e.g., WAS, LSA, word2vec, universal sentence encoder;][]{SteyEtal04, LandEtal98, MikoEtal13, CerEtal18}, here we use latent dirichlet allocation (LDA)-based topic models for a few reasons.  First, topic models capture the \textit{essence} of a text passage devoid of the specific set and order of words used.  This was an important feature of our model since different people may accurately recall a scene using very different language. Second, words can mean different things in different contexts (e.g. ``bat" may be the act of hitting a baseball, the object used for that action, or as a flying mammal).  Topic models are robust to this, allowing words to exist as part of multiple topics.  Last, topic models provide a straightforward means to recover the weights for the particular words comprising a topic, enabling easy interpretation of an event's contents (e.g. Fig.~\ref{fig:topics}). Other models such as Google's Universal Sentence Encoder offer a context-sensitive encoding of text passages, but the encoding space is complex and non-linear, and thus recovering the original words used to fit the model is not straightforward. However, it's worth pointing out that our framework is divorced from the particular choice of language model. Moreover, many of the aspects of our framework could be swapped out for other choices. For example, the language model, the timeseries segmentation model and the video-recall matching function could all be customized for the particular problem. Indeed for some problems, recovery of the particular recall words may not be necessary, and thus other text-modeling approaches (such as universal sentence encoder) may be preferable. Future work will explore the influence of particular model choices on the framework's efficacy.

In extending classical free recall analyses to our naturalistic memory framework, we recovered two patterns of recall dynamics central to list-learning studies: a heightened probability of initiating recall with the first presented ``item" (in our case, video events; Fig.~\ref{fig:list-learning}A) and a strong bias toward transitioning from recalling a given event to recalling the one immediately following it (Fig.~\ref{fig:list-learning}B).  However, equally noteworthy are the typical free recall results \textit{not} recovered in these analyses, as each highlights a fundamental difference between the list-learning paradigm and naturalistic memory paradigms like the one employed in the present study.  The most noticeable departure from hallmark free recall dynamics in these findings is the apparent lack of a serial position effect in Figure \ref{fig:list-learning}C, which instead shows greater and lesser recall probabilities for events distributed across the video.  Stimuli in free recall experiments most often comprise lists of simple, common words, presented to participants in a random order. \citep[In fact, numerous word pools have been developed based on these criteria; e.g.,][]{FrieEtal82}.  These stimulus qualities enable two assumptions that are central to word list analyses, but frequently do not hold for real-world experiences.  First, researchers conducting list-learning studies may assume that the content at each presentation index is essentially equal, and does not possess attributes that would render it, on average, more or less memorable than others.  Such is rarely the case with real-world experiences or experiments meant to approximate them, and the effects of both intrinsic and observer-dependent factors on stimulus memorability are well established \citep[for review see][]{ChunTurk07, ByliEtal15, TyngEtal17}.  Second, the random ordering of list items ensures that (across participants, on average) there is no relationship between the thematic similarity of individual stimuli and their presentation positions---in other words, two successively presented items are no more likely to be highly semantically similar than they are to be highly dissimilar.  In most cases, the exact opposite is true of real-world episodes.  Our internal thoughts, our actions, and the physical state of the world around us all tend to follow a direct, causal progression.  As a result, each moment of our experience tends to be inherently more similar to surrounding moments than to those in the distant past or future.  Memory literature has termed this strong temporal autocorrelation ``context," and in various media that depict real-world events (e.g., movies or written stories), we recognize it as a \textit{narrative structure}.  While a random word list (by definition) has no such structure, the logical progression between ideas and actions in a naturalistic stimulus prompts the rememberer to recount presented events in order, starting with the beginning.  This tendency is reflected in our findings' second departure from typical free recall dynamics: a lack of increased probability of first recall for end-of-sequence events (Fig.~\ref{fig:list-learning}A).

Because they disregard presentation order-dependent variability in the stimulus content, analyses such as those in Figure \ref{fig:list-learning} enable a more sensitive analysis of presentation order-dependent temporal dynamics in free recall. Yet by the same token, they paint a wholly incomplete picture of memory for naturalistic episodes.  In an attempt to address this shortcoming, we have developed a framework in the present study that characterizes the explicit semantic content of the stimulus and subsequent recalls.  However, sensitivity to stimulus and recall content introduces a new challenge: distinguishing between levels of recall quality for a stimulus (e.g., an event) that is considered to have been ``remembered."  When modeling memory in an experimental setting, recall quality for individual events is often cast as binary (e.g., a given list item was simply either remembered or not remembered).  Various models of memory (e.g., \citealp{Yone02}) attempt to improve upon this by including confidence ratings, rendering this binary judgement instead categorical.  To better evaluate naturalistic memory quality, we introduce a continuous metric (\textit{precision}), which reflects the level of completeness of a participant's recall for a feature-rich experience.  Additionally, recall quality for a single event is typically assessed independently from that for all other events (e.g., it is difficult to ``compare" a participant's binary recall success for list item 1 to that of list item 10).  The second novel metric we introduce (\textit{distinctiveness}) is based on analyzing of the correlational structure of an individual's full set of recall events, and reflects the specificity of their memory for a single experienced event.  We find that both of these metrics relate to the overall number of video events participants successfully recalled, and that our precision metric additionally relates to \cite{ChenEtal17}'s hand-annotated memory memory scores.  Though we do not find participants' average recall distinctiveness related to the hand-annotated memory scores, this is not entirely surprising given the divergence of behavior they capture.  In hand-scoring each participant's verbal recall for each of 50 (manually-deliniated) scenes, ``[a] scene was counted as recalled if the participant described any part of the scene" \citep{ChenEtal17}.  In other words, both an extensive description of a scene's content and a brief mention of some subset of its content were (binarily) considered equally successful recalls.  By contrast, we identify the event structure in participants' recalls in an unsupervised manner, independent of the video event-timeseries, prior to mapping between video and recall content.  Our HMM-based event-segmentation produces boundaries between timepoints where the topic proportions shift in a substantial way, and because a small handful of words is unlikely to contribute significantly to the topic proportions for any sliding window, such brief scene descriptions will most often not begat a sufficiently large shift in the resulting topic proportions for the HMM to identify an event boundary.  Instead, they will be grouped with a neighboring event, consequently lowering that event's distinctiveness score and by extension, the participant's overall distinctiveness score.  This is in essence the qualitative difference between distinctive and indistinctive recall, and reflects the comparison shown in Figure \ref{fig:distinctiveness-detail}C.  Intriguingly, prior studies show that pattern separation, or the ability to cleanly discriminate between similar experiences, is impaired in many cognitive disorders as well as natural aging \citep{StarEtal10, YassEtal11c, YassStar11b}.  Future work might explore whether and how these metrics compare between cognitively impoverished groups and healthy controls.

In the analyses outlined in Figure \ref{fig:brainz}, we identified two networks of brain regions whose responses during video viewing were consistent with the temporal structure of the video and recall topic trajectories, respectively. The network identified by the video trajectory analysis included the ventromedial prefrontal cortex, left anterior temporal lobe, superior parietal and dorsal anterior cingulate cortex. The network from the video-recall trajectory analysis also included the ventromedial prefrontal and superior parietal cortices, in addition to the posterior medial cortex (PMC) and the inferior temporal regions. Notably, \cite{ChenEtal17} also observed the PMC in a number of analyses including one that searched for regions whose activity patterns during encoding were reinstated during free recall. The PMC has been consistently identified in studies involving stimuli with meaningfully structured events \cite{CohnRang17}. Further, the PMC is part of the "posterior medial" system, a network of brain regions thought to represent situation models ~\cite{ZackEtal07} in support of memory, spatial navigation and social cognition ~\citep{RangRitc12}. Given that we constructed our video-recall searchlight model to capture temporal structure in the episode's semantic content (and how one's later recall aligns with that structure), we speculate that the PMC may play a role in constructing mnemonic events from meaningfully structured experiences.

Decoding the associated significance maps with \texttt{Neurosynth} revealed two intriguing results.  First, the top 10 terms returned for the video-driven searchlight significance map were centered around themes of language and semantic meaning (Fig.~\ref{fig:neurosynth}A).  In other words, the voxels identified as more reflective of the video's temporal structure (i.e., voxels with lower permutation correction-derived $p$-values), as defined by our model, were most likely to be reported as active in studies focused on the the neural underpinnings of semantic processing.  This finding is interesting, as our model specifically captures the temporal structure of the video's \textit{semantic} content (e.g., as opposed to that of the visual, auditory, or affective content).  This suggests that the network of structures displayed in Figure \ref{fig:brainz}C may play a roll in processing the evolving semantic structure of ongoing experiences.

Our second searchlight analysis identified a largely separate network of regions (Fig.~\ref{fig:brainz}D) whose patterns of activity as participants viewed the video reflected the idiosyncratic structure of each individual's later recall.  Decoding the associated significance map yielded a set of terms that primarily reflected names of specific structural regions (such as ``thalamus," ``anterior insula," ``anterior cingulate" and ``inferior frontal"; Fig.~\ref{fig:neurosynth}B).  Interestingly, these regions share membership in a common, large-scale functional network (termed the ``salience network") involved in detecting and processing affective cues.  In particular, the latter three regions have been implicated in functions relevant to assigning personal meaning to an experience, including: ascribing subjective value to raw, sensory input \citep{MedfCrit10}; modulating semantic and phonological processing in response to personally salient stimuli \citep{KellEtal07b}; and directing and reallocating attention and working memory resources towards the most relevant stimuli \citep{MenoUddi10}.  This suggests that the network of structures displayed in Figure \ref{fig:brainz}D may be play a roll in transforming and restructuring ongoing experiences through the lens of one's own personal values as they are encoded in memory.

>>>>

\begin{abstract}
Our moment-by-moment subject experience reflects external sensory information from each moment, along with additional information from our past that we carry with us into that moment.  The blend of memories, knowledge, emotions, goals, and other internal perceptual and mental states that color our subjective experience provides a \textit{context} for interpreting new information and conceptually linking what is happening now with our prior experiences.  Because this contextual information is often person-specific, the subjective experience that each person encodes into their memory is often idiosyncratic, even for shared experiences and sensory perspectives.  We sought to study which aspects of a shared naturalistic experience were preserved or distorted, and how those distortions compared across individuals.  To this end, we developed a geometric framework for mathematically characterizing the subjective conceptual content of dynamic naturalistic experiences.  We model experiences as \textit{trajectories} through word embedding spaces whose coordinates reflect the universe of thoughts under consideration.  We also demonstrate how \textit{memories} may also be modeled as trajectories through the same spaces.  According to this view, encoding an experience into memory entails geometrically distorting or transforming the original experience's trajectory.  This translates qualitative neuropsychological questions about how we remember naturalistic experiences into quantitive geometric questions about the spatial configurations of trajectory shapes.  We applied our framework to data collected as participants watched and verbally recounted a television episode while undergoing functional neuroimaging.  We found that the trajectories of participants' recountings of the episode nearly all captured the coarse spatial properties of the original episode's trajectory (i.e., the essential plot points), but participants differed in their memory for fine details.  We also identified a network of brain structures that were sensitive to the shape of the episode's trajectory through word embedding space, and an overlapping network that predicted, at the time of encoding, how people would distort (transform) the episode's trajectory when they recounted the episode later.  Our work provides insights into how our brains distort and transform our ongoing experiences when we encode them into episodic memories.
\end{abstract}

@@

Statistical models of memory studies often treat recall success as binary (in other words, an item either was or was not recalled), or occasionally categorical \citep[e.g., to distinguish familiarity from recollection;][]{YoneEtal02}.  Such approaches are tenable in classical list-learning or recognition memory paradigms, as the presented stimuli tend to be very simple (e.g., a sequence of individual words or items).  However, memory for naturalistic experiences is much more nuanced.  For example, certain aspects of an experience might be correctly remembered at varying levels of detail, or distorted, or forgotten entirely.  Further, each remembering is itself a richly structured phenomenon.  Our framework produces a content-based model of individual video and recall events by projecting the dynamic content of the video and participants' recalls into a shared topic space.  This allows for direct, quantitative comparisons between all stimulus and recall events, as well as between the recall events themselves.  Leveraging these content-based models of the stimulus/recall events, we developed two novel, \textit{continuous} metrics for analyzing naturalistic memory:~\textit{precision} and \textit{distinctiveness}.  Precision is intended to capture the ``completeness'' of recall, or how fully the presented content was recapitulated in memory.  We define a recall event's precision as the maximum correlation between the topic proportions of that recall event and any video event (Fig.~\ref{fig:precision-distinctiveness}).  A second novel metric we introduce here is \textit{distinctiveness}, which is intended to capture the ``specificity'' of recall.  In other words, distinctiveness quantifies the extent to which a given recalled event reflects the most similar presented event moreso than it does other presented events.  To compute a recall event's distinctiveness, we first identify the video event to which its topic vector is most strongly correlated.  We then define distinctiveness as one minus the average correlation between the given recall event and all \textit{other} video events.  In addition to individual events, one may also use these metrics to describe each participant's overall performance by averaging across a participant's event-wise precision or distinctiveness scores.  Participants whose recall events are more veridical descriptions of what happened in the video event will presumably have higher precision scores. We find that, across participants, higher precision scores are positively correlated with both hand-annotated memory performance (as collected by Chen et al., 2017; Pearson's $r(15) = 0.60, p = 0.010$) and the number of video events successfully remembered, as determined by our model (Pearson's $r(15) = 0.90, p < 0.001$).  We also hypothesized that participants who recounted events in a more distinctive way would display better overall memory.  We find that participants' distinctiveness scores were positively
correlated with our model's estimated number of recall events (Pearson's $r(15) = 0.55, p = 0.021$).  However, we found no evidence that distinctiveness scores were correlated with hand-annotated memory performance (Pearson's $r(15) = 0.28, p = 0.275$).  We elaborate on this potential discrepency in the \textit{Discussion} section.

@@

Further intuition for the behaviors captured by these two metrics may be gained by directly examining the content of the video and recalls our framework models.  In Figure \ref{fig:precision-detail}, we contrast recalls for the same video event (event 22) from two participants: one with a high precision score (P17), the other with a low precision score (P6).  From the HMM-identified event boundaries, we recovered the set of annotations describing the content of an example video event (Fig.~\ref{fig:precision-detail}B), and divided them into different color-coded sections for each action or feature described.  We then similarly recovered the set of sentences comprising the corresponding recall event for each of the two example participants.  Because the recall sliding windows overlap heavily, and each recall event spans multiple recall timepoints (i.e., windows), we have stripped any sentences from the beginning and end that describe earlier or later video events for the sake of readability.  In other words, Fig.~\ref{fig:precision-detail}C shows a subset of the full recall event text, comprising sentences between the first and last descriptions of content from the example video event.  We then colored all words describing actions and features coded in panel B by their corresponding color.  Visual comparison of these example transcripts reveals that the more precise recall captures more of the video event's content, and with more detail.

Figure \ref{fig:distinctiveness-detail} similarly contrasts two example participants' recalls for a common video event (event 19) to illustrate the tangible differences between high and low distinctiveness scores.  Here, we have extracted the full set of sentences comprising the most distinctive recall event (P13) and least distinctive recall event (P11) matched to the example video event (Fig.~\ref{fig:distinctiveness-detail}C).  We also extracted the annotations for the example video event, as well as those from each other video event whose content the example participants' single recall events described (Fig.~\ref{fig:distinctiveness-detail}B).  We then shaded the annotation text for each video event with a different color, and shaded each word of the example participants' recall text by the color of the video event it describes.  The majority of the most distinctive recall event text describes video event 19's content, with the first five and last one sentence describing the video events immediately preceding and succeeding the current one, respectively.  In contrast, the least distinctive recall of video event 19 blends the content from five separate video events, does not transition between them in order, and often combines descriptions of two video events' content in the same sentence.

@@

In addition to the more ``holistic'' measure of memory described in the previous section, our framework also affords the ability to drill down to individual words and quantify how each word relates to the  memorability of each event. The results displayed in Figures \ref{fig:list-learning}C and \ref{fig:precision-detail}A suggest that certain events were remembered better than others.  Given this, we next asked asked whether the events were generally remembered well or poorly tended to reflect particular content.  Because our analysis framework projects the dynamic video content and participants' recalls into a shared space, and because the dimensions of that space represent topics (which are, in turn, sets of weights over known words in the vocabulary), we are able to recover the weighted combination of words that make up any point (i.e., topic vector) in this space.  We first computed the average precision with which participants recalled each of the 30 video events (Fig.~\ref{fig:topics}A; note that this result is analogous to a serial position curve created from our continuous recall quality metric).  We then computed a weighted average of the topic vectors for each video event, where the weights reflected how reliably each event was recalled.  To visualize the result, we created a ``wordle'' image~\citep{MuelEtal18} where words weighted more heavily by better-remembered topics appear in a larger font (Fig.~\ref{fig:topics}B, green box).  Across the full video, content that reflected topics necessary to convey the central focus of the video (e.g., the names of the two main characters, ``Sherlock'' and ``John,'' and the address of a major recurring location, ``221B Baker Street") were best remembered.  An analogous analysis revealed which themes were poorly remembered.  Here in computing the weighted average over events' topic vectors, we weighted each event in \textit{inverse} proportion to how well it was remembered (Fig.~\ref{fig:topics}B, red box).  The least well-remembered video content reflected information not necessary to later convey a general summary of the video, such as the proper names of relatively minor characters (e.g., ``Mike," ``Molly," and ``Lestrade") and locations (e.g., ``St. Bartholomew's Hospital").
%, as well as the brief, animated clip participants viewed at the beginning of each of the two scan session (involving ``singing" ``cartoon" characters).

@@

A similar result emerged from assessing the topic vectors for individual video and recall events (Fig.~\ref{fig:topics}C).  Here, for each of the three best- and worst-remembered video events, we have constructed two wordles: one from the original video event's topic vector (left) and a second from the average recall topic vector for that event (right).  The three best-remembered events (circled in green) correspond to scenes integral to the central plot-line: a mysterious figure spying on John in a phone booth; John meeting Sherlock at Baker St. to discuss the murders; and Sherlock laying a trap to catch the killer.  Meanwhile, the three worst-remembered events (circled in red) reflect scenes that are non-essential to summarizing the narrative's structure: the video of singing cartoon characters participants viewed in an introductory clip prior to the main episode; John asking Molly about Sherlock's habit of over-analyzing people; and Sherlock noticing evidence of Anderson's and Donovan's affair.

@@

Word embedding models are a rapidly growing area of machine learning research.  Early approaches including latent
  semantic analysis~\citep{LandDuma97} use word co-occurrence statistics (i.e., how often pairs of words occur in the same documents contained in the corpus) to derive a unique feature vector for each word.  The feature vectors are constructed so that words that co-occur more frequently have feature vectors that are closer (in Euclidean distance).  Related approaches, such as latent dirichlet
  allocation~\citep{BleiEtal03} attempt to explicitly model the underlying causes of word co-occurrences by automatically identifying the set of themes or topics reflected across the documents in the corpus.  More recent work on these types of semantic models, including word2vec~\citep{MikoEtal13}, the Universal Sentence Encoder~\citep{CerEtal18}, GPT-2~\citep{RadfEtal19}, and GTP-3~\citep{BrowEtal20} use deep neural networks to attempt to identify the deeper conceptual representations underlying each word.  Despite the growing popularity of more sophisticated deep learning-based embedding models, here we leverage latent dirichlet allocation (i.e., topic modelling) to embed video and recall text.  This decision was motivated by several factors.  First, topic models capture the \textit{essence} of a text passage devoid of the specific set and order of words used.  This was an important feature of our model since different people may accurately recall a scene using very different language. Second, words can mean different things in different contexts (e.g. ``bat'' may be the act of hitting a baseball, the object used for that action, or as a flying mammal).  Topic models are robust to this, allowing words to exist as part of multiple topics.  Last, topic models provide a straightforward means of recovering the weights for the particular words comprising a topic, enabling straightforward interpretation of an event's contents (e.g. Fig.~\ref{fig:topics}). Other models such as the Universal Sentence Encoder, GPT-2, and GPT-3 offer context-sensitive encoding of text passages, but the encoding space is complex and non-linear, and thus recovering the original words used to fit the model is not straightforward. However, it is worth pointing out that our general framework is divorced from the particular choice of language model. Moreover, many of the aspects of our framework could be swapped out for other choices. For example, the language model, the timeseries segmentation model and the video-recall matching function could all be customized to suit a particular question space or application. Indeed for some questions, recovery of the particular words used to describe a memory may not be necessary, and thus other text-modeling approaches (including the deep learning-based embedding models described above) may be preferable. Future work will explore the influence of particular model choices on the framework's efficacy.

In extending classical free recall analyses to our naturalistic memory framework, we recovered two patterns of recall dynamics central to list-learning studies: a heightened probability of initiating recall with the first presented ``item" (in our case, video events; Fig.~\ref{fig:list-learning}A) and a strong bias toward transitioning from recalling a given event to recalling the one immediately following it (Fig.~\ref{fig:list-learning}B).  However, equally noteworthy are the typical free recall results \textit{not} recovered in these analyses, as each highlights a fundamental difference between the list-learning paradigm and naturalistic memory paradigms like the one employed in the present study.  The most noticeable departure from hallmark free recall dynamics in these findings is the apparent lack of a serial position effect in Figure \ref{fig:list-learning}C, which instead shows greater and lesser recall probabilities for events distributed across the video.  Stimuli in free recall experiments most often comprise lists of simple, common words, presented to participants in a random order. \citep[In fact, numerous word pools have been developed based on these criteria; e.g.,][]{FrieEtal82}.  These stimulus qualities enable two assumptions that are central to word list analyses, but frequently do not hold for real-world experiences.  First, researchers conducting list-learning studies may assume that the content at each presentation index is essentially equal, and does not possess attributes that would render it, on average, more or less memorable than others.  Such is rarely the case with real-world experiences or experiments meant to approximate them, and the effects of both intrinsic and observer-dependent factors on stimulus memorability are well established \citep[for review see][]{ChunTurk07, ByliEtal15, TyngEtal17}.  Second, the random ordering of list items ensures that (across participants, on average) there is no relationship between the thematic similarity of individual stimuli and their presentation positions---in other words, two successively presented items are no more likely to be highly semantically similar than they are to be highly dissimilar.  In most cases, the exact opposite is true of real-world episodes.  Our internal thoughts, our actions, and the physical state of the world around us all tend to follow a direct (often causal) progression.  As a result, each moment of our experience tends to be inherently more similar to surrounding moments than to those in the distant past or future.  Memory literature has termed this strong temporal autocorrelation ``context," and in various media that depict real-world events (e.g., movies or written stories), we recognize it as a \textit{narrative structure}.  While a random word list (by definition) has no such structure, the logical progression between ideas and actions in a naturalistic stimulus prompts the rememberer to recount presented events in order, starting with the beginning.  This tendency is reflected in our findings' second departure from typical free recall dynamics: a lack of increased probability of first recall for end-of-sequence events (Fig.~\ref{fig:list-learning}A).

Because they disregard presentation order-dependent variability in the stimulus content, analyses such as those in Figure \ref{fig:list-learning} enable a more sensitive analysis of presentation order-dependent temporal dynamics in free recall. Yet by the same token, they paint a wholly incomplete picture of memory for naturalistic episodes.  In an attempt to address this shortcoming, we have developed a framework in the present study that characterizes the explicit semantic content of the stimulus and subsequent recalls.  However, sensitivity to stimulus and recall content introduces a new challenge: distinguishing between levels of recall quality for a stimulus (e.g., an event) that is considered to have been ``remembered."  When modeling memory in an experimental setting, recall quality for individual events is often cast as binary (e.g., a given list item was simply either remembered or not remembered).  Various models of memory (e.g., \citealp{Yone02}) attempt to improve upon this by including confidence ratings, rendering this binary judgement instead categorical.  To better evaluate naturalistic memory quality, we introduce a continuous metric (\textit{precision}), which reflects the level of completeness of a participant's recall for a feature-rich experience.  Additionally, recall quality for a single event is typically assessed independently from that for all other events (e.g., it is difficult to ``compare" a participant's binary recall success for list item 1 to that of list item 10).  The second novel metric we introduce (\textit{distinctiveness}) is based on analyzing of the correlational structure of an individual's full set of recall events, and reflects the specificity of their memory for a single experienced event.  We find that both of these metrics relate to the overall number of video events participants successfully recalled, and that our precision metric additionally relates to \cite{ChenEtal17}'s hand-annotated memory memory scores.

We did not find evidence that participants' average recall distinctiveness was related to their hand-annotated memory scores computed by~\cite{ChenEtal17}.  One possible explanation is that, in hand-scoring each participant's verbal recall for each of 50 (manually-deliniated) scenes, ``[a] scene was counted as recalled if the participant described any part of the scene''~\citep{ChenEtal17}.  In other words, both an extensive description of a scene's content and a brief mention of some subset of its content were (binarily) considered equally successful recalls.  By contrast, we identify the event structure in participants' recalls in an unsupervised manner, independent of the video event-timeseries, prior to mapping between video and recall content.  Our HMM-based event-segmentation produces boundaries between timepoints where the topic proportions shift in a substantial way, and because a small handful of words is unlikely to contribute significantly to the topic proportions for any sliding window, such brief scene descriptions will most often not result in a sufficiently large shift in the resulting topic proportions for the HMM to identify an event boundary.  Instead, they will be grouped with a neighboring event, consequently lowering that event's distinctiveness score and by extension, the participant's overall distinctiveness score.  This is in essence the qualitative difference between distinctive and indistinctive recall, and reflects the comparison shown in Figure \ref{fig:distinctiveness-detail}C.  Intriguingly, prior studies show that pattern separation, or the ability to cleanly discriminate between similar experiences, is impaired in many cognitive disorders as well as natural aging \citep{StarEtal10, YassEtal11c, YassStar11b}.  Future work might explore whether and how these metrics compare between cognitively impoverished groups and healthy controls.

In the analyses outlined in Figure \ref{fig:brainz}, we identified two networks of brain regions whose responses during video viewing were consistent with the temporal structure of the video and recall topic trajectories, respectively. The network identified by the video trajectory analysis included the ventromedial prefrontal cortex, left anterior temporal lobe, superior parietal and dorsal anterior cingulate cortex. The network from the video-recall trajectory analysis also included the ventromedial prefrontal and superior parietal cortices, in addition to the posterior medial cortex (PMC) and the inferior temporal regions. Notably, \cite{ChenEtal17} also observed the PMC in a number of analyses including one that searched for regions whose activity patterns during encoding were reinstated during free recall. The PMC has been consistently identified in studies involving stimuli with meaningfully structured events~\citep{CohnRang17}. Further, the PMC is part of the "posterior medial" system, a network of brain regions thought to represent situation models~\citep{ZackEtal07} in support of memory, spatial navigation and social cognition~\citep{RangRitc12}. Given that we constructed our video-recall searchlight model to capture temporal structure in the episode's semantic content (and how one's later recall aligns with that structure), we speculate that the PMC may play a role in constructing mnemonic events from meaningfully structured experiences.

Decoding the associated significance maps with \texttt{Neurosynth} revealed two intriguing results.  First, the top 10 terms returned for the video-driven searchlight significance map were centered around themes of language and semantic meaning (Fig.~\ref{fig:neurosynth}A).  In other words, the voxels identified as more reflective of the video content's temporal structure (i.e., voxels with lower permutation correction-derived $p$-values), as defined by our model, were most likely to be reported as active in studies focused on the the neural underpinnings of semantic processing.  This finding is interesting, as our model specifically captures the temporal structure of the video's \textit{semantic} content (e.g., as opposed to that of the visual, auditory, or affective content).  This suggests that the network of structures displayed in Figure \ref{fig:brainz}C may play a roll in processing the evolving semantic content of ongoing experiences.
